### 2024-04-02 other/datasheet/datasheet_template.qmd
Initial commit

---
title: "Datasheet for 'A dataset'"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "02 April 2024"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - TBD
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - TBD
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - TBD
4. *Any other comments?*
    - TBD

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- TBD
2. *How many instances are there in total (of each type, if appropriate)?*
	- TBD
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- TBD
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- TBD
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- TBD
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- TBD
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- TBD
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- TBD
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- TBD
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- TBD
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- TBD
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- TBD
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- TBD
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- TBD
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- TBD
16. *Any other comments?*
	- TBD

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References

### 2024-04-02 paper/paper.qmd
Initial commit

---
title: "Are Greener Spaces, Safer Places?"
subtitle: "Investigating the Link Between Green Spaces and Crime in Toronto"
author: 
  - Jena Shah
thanks: "Code and data are available at: https://github.com/shahjena/greenspace_safeplace.git."
date: today
date-format: long
abstract: "Abstract TBD."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

TBD

# Data {#sec-data}

TBD

# Model

TBD

## Model set-up

TBD

## Model justification

TBD

# Results

TBD

# Discussion

## First discussion point {#sec-first-point}

TBD

## Second discussion point

TBD

## Third discussion point

TBD

## Weaknesses and next steps

TBD

\newpage

\appendix

# Appendix {-}

# Datasheet

TBD

# Model details {#sec-model-details}

# Posterior predictive check

TBD

# Diagnostics

TBD

\newpage

# References

@citeR

### 2024-04-02 scripts/00-simulate_data.R

@@ -1,0 +1,12 @@
+#### Preamble ####
+# Purpose: Simulates data.
+# Author: Jena Shah
+# Date: 02 April 2024
+# Contact: jena.shah@mail.utoronto.ca
+# License: MIT
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Simulate data ####

### 2024-04-02 scripts/01-download_data.R

@@ -1,0 +1,21 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data.
+# Author: Jena Shah
+# Date: 02 April 2024
+# Contact: jena.shah@mail.utoronto.ca
+# License: MIT
+
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(tidyverse)
+
+#### Download data ####
+# Download data from opendatatoronto (1) wellbeing-toronto-environment (2) neighbourhood-crime-rates.
+# Download (1) as the_green_spaces and (2) as the_crime_rates.
+
+#### Save data ####
+write_csv(the_crime_rates, "inputs/data/crime_rates.csv") 
+write_csv(the_green_spaces, "inputs/data/green_spaces.csv")
+
+         

### 2024-04-02 scripts/02-data_cleaning.R

@@ -1,0 +1,15 @@
+#### Preamble ####
+# Purpose: Cleans the green spaces and crime rates data and merges them.
+# Author: Jena Shah
+# Date: 02 April 2024
+# Contact: jena.shah@mail.utoronto.ca
+# License: MIT
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Clean data ####
+raw_data <- read_csv("inputs/data/plane_data.csv")
+
+#### Save data ####
+write_csv(cleaned_data, "outputs/data/analysis_data.csv")

### 2024-04-02 scripts/03-test_data.R

@@ -1,0 +1,12 @@
+#### Preamble ####
+# Purpose: Tests the cleaned data.
+# Author: Jena Shah
+# Date: 02 April 2024
+# Contact: jena.shah@mail.utoronto.ca
+# License: MIT
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Test data ####

### 2024-04-02 scripts/04-model.R

@@ -1,0 +1,25 @@
+#### Preamble ####
+# Purpose: Models data.
+# Author: Jena Shah
+# Date: 02 April 2024
+# Contact: jena.shah@mail.utoronto.ca
+# License: MIT
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read data ####
+analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+
+### Model data ####
+# Model here assigned to my_model
+
+#### Save model ####
+saveRDS(
+  my_model,
+  file = "models/my_model.rds"
+)
+
+

### 2024-04-19 other/datasheet/datasheet_template.qmd
Commit

@@
 ---
-title: "Datasheet for 'A dataset'"
-subtitle: "My subtitle if needed"
+title: "Datasheet for 'Are Greener Spaces, Safer Places?'"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
-date: "`r Sys.time()`"
-date-format: "02 April 2024"
+  - Jena Shah
+thanks: "Code and data are available at: https://github.com/shahjena/greenspace_safeplace.git."
+date: today
+date-format: long
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."

### 2024-04-19 paper/paper.qmd
Commit

@@
 ---
-title: "Are Greener Spaces, Safer Places?"
-subtitle: "Investigating the Link Between Green Spaces and Crime in Toronto"
+title: "Are Greener Spaces Less Safe Places?"
+subtitle: "Exploring the Surprising Relationship Between Urban Green Spaces and Crime Rates in Toronto"
 author: 
@@
 thanks: "Code and data are available at: https://github.com/shahjena/greenspace_safeplace.git."
+toc: true
 date: today
 date-format: long
-abstract: "Abstract TBD."
+abstract: "This study examines the relationship between green spaces and crime rates in Toronto's neighborhoods using gamma regression models.
+The analysis reveals a positive association between the amount of green spaces and crime rates, indicating that neighborhoods with more green spaces tend to have higher crime rates.
+This finding emphasizes the relation between urban green spaces and public safety, suggesting that the presence of green spaces may present challenges for crime prevention and urban planning.
+Understanding these dynamics can help the government and urban planners in their efforts to create safer and more sustainable cities."
 format: pdf
@@
 bibliography: references.bib
+editor: 
+  markdown: 
+    wrap: 72
 ---
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+library(tidyverse)
+library(marginaleffects)
+library(arrow)
+library(rstanarm)
+library(ggplot2)
+```
+\newpage
 
@@
 
-TBD
+In urban environments, the presence of green spaces has long been
+associated with a variety of benefits: from improved mental health to
+enhanced community cohesion. However, the relationship between green
+spaces and crime rates is a complex one and has been a subject of debate
+in urban planning and crime prevention. There are plenty of studies
+suggesting that green spaces may act as a deterrent to crime through
+promoting community interaction and physical activity. However, there
+are also others who raise concerns that these areas may attract criminal
+activity. Understanding this relationship is crucial for creating safer
+and more livable cities.
 
-# Data {#sec-data}
+National Geographic has summarized studies indicating that urban areas
+with more green spaces tend to have lower crime rates. For example, in
+Flint, Michigan - a city that experienced a decrease in crime rates from
+2012 to 2022 - the implementation of green spaces was correlated with
+this positive change. Similarly, places like Ohio and Philadelphia have
+seen recommendations to increase green spaces in areas with high crime
+rates as a potential strategy to reduce crime [@burtka2023]. Not only
+that, but there is further research backing up the results discussed by
+National Geographic. A research paper from the Multidisciplinary Digital
+Publishing Institute (MDPI) conducted a comprehensive literature review
+to investigate the impact of green spaces on violent crime in urban
+settings. Their review included over 30000 potential paper titles and
+they ultimately selected 45 papers for inclusion. After their analysis,
+their findings and results suggested that green spaces play a
+significant role in reducing violent crime. The mechanisms proposed for
+this impact include social interaction, social recreation, community
+perception, nature-induced stress reduction, climate change, and many
+more [@shepley2019].
 
-TBD
+Unlike the two studies mentioned, this paper will represent the controversial side's contribution to the
+current understanding of urban dynamics by focusing on the relationship
+between the prevalence of green spaces and crime rates within Toronto's
+urban neighbourhoods. By utilizing a rich and expansive dataset along
+with a sophisticated statistical model, we conduct a thorough analysis
+to unravel the complexities of this relationship. In doing so, the
+findings not only focus on the interconnections between green spaces and
+crime rates, but also offer insights that can significantly inform the
+formulation of urban planning strategies for crime prevention. The
+estimand for this study is to quantify the effect of an increase in
+green spaces on crime rates in urban neighbourhoods - while controlling
+for relevant factors.
+
+The paper is structured into four main sections. First, I provide a
+detailed overview of the data utilized in this study, including its
+sources and any pre-processing/cleaning steps. I also describe the
+methodology used for the analysis, including the statistical figures
+made to explore the relationship between green spaces and crime rates.
+After that, I present the modeling approach used in this study -
+detailing the statistical techniques applied to analyze the data. This
+section includes a discussion of the model setup and justification for
+the chosen modeling approach. Following that, I present the results of
+my analysis, highlighting the key findings regarding the relationship
+between green spaces and crime rates in urban neighbourhoods. I use
+visualizations such as graphs and tables to illustrate these results
+effectively. Finally, the implications of my findings are discussed for
+urban planning and crime prevention. I explore how these results can
+better the policies which are aimed at creating safer and more
+sustainable urban environments. I also take a look at the weaknesses of
+my study which could bias the results. Throughout the paper, I utilized
+the R programming language and its packages which are cited in the
+references.
+
+```{r}
+#| label: fig-one
+#| fig-cap: Distribution of Green.Spaces and TOTAL_CRIMES.
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: true
+cleaned_data <- read.csv("C:/Users/shahj/OneDrive/Documents/greencrimes/data/cleaned/cleaned_data.csv")
+total_2023 <- cleaned_data$TOTAL_CRIMES
+green_spaces_2023 <- cleaned_data$Green.Spaces
+
+par(mfrow = c(1, 2))
+
+hist(total_2023, main = "Total Crimes per 100000, 2023", xlab = "Rate")
+hist(green_spaces_2023, main = "Green Spaces 2023", xlab = "Percentage")
+
+par(mfrow = c(1, 1))
+
+```
+
+# Data
+
+## Data Collection and Cleaning
+
+The datasets for this project were obtained from the City of Toronto's
+open data portal. There were many other sources that could provide
+similar data but they were much older, hence it was more appropriate to
+use the data from Open Data Toronto's portal. The first dataset,
+"Neighbourhood Crime Rates," contains all neighbourhoods in Toronto and gives the rates (out of 100000) of nine different types of crime from the years 2014 to 2023.
+Example of crimes include, Assault, Auto Theft, Break and Enter, Robbery, Theft Over \$5000, and more.
+The second dataset, "Wellbeing Toronto Environment," contains information on green spaces and environmental factors. This dataset was provided by Toronto Public
+Health, Toronto Parks, Forestry and Recreation, the Federal Ministry of
+the Environment, and Toronto Water. It includes data on Green Rebate
+Programs, Pollutant Carcinogenic TEP Score, Pollutant Non-Carcinogenic
+TEP Score, and Pollutants Released to Air, among other variables.
+
+The datasets were combined using the neighbourhood ID (HOOD_ID) and
+filtered for the variables of interest, which include the neighbourhood
+ID, neighbourhood name (AREA_NAME), and crime rates for Assault, Auto
+Theft, Break and Enter, Robbery, and Theft Over for the year 2023.
+Additionally, the dataset includes information on Green Spaces for each
+neighbourhood. Since the data of crime across Toronto did not provide a
+total amount of crimes per neighbourhood and only focused on the
+different types of crime and their rates, I created a new variable which
+combined all the different crimes and merged it into a variable called
+TOTAL_CRIMES which provides the total amount of crimes per 100000 people
+in a neighbourhood.
+
+The decision to focus on the year 2023 out of all the years from 2014 to
+2023 was made to use the most recent data available. In the crimes data,
+there were a total of 9 types of crimes used but this paper focuses on
+using the top 5 crimes in Toronto in the year of 2023 according to
+Vilkhov Law. They were selected for analysis and include Assault, Auto
+Theft, Break and Enter, Robbery, and Theft Over [@vilkhovlaw]. Other
+years and additional crimes were available in the raw crime data, but
+for the purposes of this study, only the year 2023 and the selected top
+5 crimes were considered.
+
+Overall, the data collection and pre-processing steps involved
+obtaining, combining, and filtering the datasets to create a
+consolidated dataset for analysis. These steps ensured that the data
+used in the study were relevant, accurate, and suitable for
+investigating the relationship between green spaces and crime rates in
+Toronto's neighbourhoods.
+
+## Data Variables
+
+The dataset for this study comprises four key variables that are central
+to the analysis. These variables are properly descibed below.
+
+1.  HOOD_ID (Neighbourhood ID): This variable serves as a unique
+    identifier for each neighborhood in Toronto, allowing us to
+    differentiate between different areas in our analysis.
+
+2.  AREA_NAME (Neighbourhood Name): This variable provides the name of
+    each neighborhood, providing context and a more intuitive
+    understanding of the data.
+
+3.  Green.Spaces (Amount of Green Spaces): This variable quantifies the
+    amount of green spaces in each neighborhood, measured in square
+    kilometers. Green spaces include parks, forests, and other areas
+    covered by vegetation.
+
+4.  TOTAL_CRIMES (Total Crimes per Neighbourhood per 100000 People):
+    This variable represents the total number of reported crimes per
+    neighborhood, standardized to a rate per 100000 people. It provides
+    a measure of the overall crime burden in each neighborhood,
+    accounting for differences in population size.
+
+These variables were selected based on their relevance to the study's
+objective of examining the relationship between green spaces and crime
+rates. The inclusion of the neighborhood ID and name allows us to
+identify and differentiate between neighborhoods, while the Green.Spaces
+variable provides a measure of the presence and extent of green spaces
+in each area. The TOTAL_CRIMES variable provides a measure of the
+overall crime rates, allowing us to assess the impact of green spaces on
+crime in different neighborhoods. @fig-one shows the different variables
+and we can see the right-skew in the crimes total, which can help us
+decide what regression model to use in the next step.
 
@@
 
-TBD
+## Model Set-up
 
-## Model set-up
+This study was done to investigate the relationship between the abundance of green
+spaces and crime rates in urban neighborhoods in Toronto by using a
+gamma regression model. This approach is suitable for analyzing count
+data with a skewed distribution, such as crime rates, where the response
+variable is non-negative and continuous.
 
-TBD
+### Variables
 
-## Model justification
+The key variables used in the model are:
 
-TBD
+-   Dependent Variable: TOTAL_CRIMES (total crimes per neighborhood per
+    100000 people).
+
+-   Independent Variable: Green.Spaces (amount of green spaces in square
+    kilometers).
+
+### Model Details
+
+The gamma regression model is specified as follows:
+
+$log(E(Y)) = \beta_0 + \beta_1 \times \mbox{Green.Spaces}$
+
+Where:
+
+-   (E(Y)) is the expected value of the dependent variable
+    (TOTAL_CRIMES).
+
+-   $\beta_0$ represents the intercept term.
+
+-   $\beta_1$ is the coefficient associated with the Green.Spaces
+    variable, indicating the effect of green spaces on crime rates.
+
+### Assumptions
+
+The gamma regression model assumes that the dependent variable follows a
+gamma distribution. We also assume that the relationship between green
+spaces and crime rates is linear, meaning that a one-unit increase
+in green spaces is associated with a constant percentage change in crime
+rates.
+
+## Model Justification
+
+The choice of a gamma regression model is appropriate for this study
+because it is specifically designed for analyzing skewed count data. We noticed the right-skew in the histograms of green spaces and total crimes in @fig-one. 
+By using a gamma regression approach, we can account for the skewed nature of crime rate data and obtain more reliable estimates of the
+relationship between green spaces and crime rates.
+
+Overall, the gamma regression model is well-suited for analyzing the
+relationship between green spaces and crime rates in Toronto's
+neighborhoods, providing a statistically interpretable
+approach to examining this important issue.
 
@@
 
-TBD
+In the results, I found that green spaces have a statistically significant relationship with crime rates in Toronto's neighborhoods.
+The gamma regression model showed that as the amount of green spaces in a neighborhood increases, the total crimes per 100000 people also tend to increase.
+Specifically, for every unit increase in green spaces (measured in square kilometers), there was a corresponding increase in total crimes.
+
+@tab-one presents the results of the explanatory model for crime rates based on green spaces.
+The intercept and coefficient for green spaces are both statistically significant, indicating that green spaces have a significant impact on crime rates.
+The model had 122 observations, and the log-likelihood was -943.227.
+The model's predictive performance was assessed using ELPD, LOOIC, WAIC, and RMSE, indicating a good fit of the model to the data.
+
+@fig-one illustrates this relationship with a scatterplot.
+It shows a concentration of data points at low to medium crime rates and low levels of green spaces.
+As the amount of green spaces increases, there are much fewer data points and they are more scattered, indicating higher crime rates in neighborhoods with more green spaces.
+The line of best fit in the scatterplot shows a nearly straight line, yet still highlighting the positive relationship between green spaces and crime rates.
+
+@fig-two enhances this by presenting a line graph that shows the uncertainty shadow of the line of best fit.
+It demonstrates that the shadow is small at low to medium crime rates and low levels of green spaces, indicating higher certainty in the relationship.
+However, as the amount of green spaces increases, the shadow expands a lot, indicating more uncertainty in the relationship.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+cleaned_par <- read_parquet("C:/Users/shahj/OneDrive/Documents/greencrimes/data/cleaned/cleaned_data.parquet") %>%
+  filter(!is.na(TOTAL_CRIMES) & TOTAL_CRIMES != 0 & !is.na(Green.Spaces) & Green.Spaces != 0)
+
+```
+
+```{r}
+#| label: fig-two
+#| fig-cap: Crime rate (out of 100000) frequency depending on the area of green spaces (square kilometers)
+#| echo: false
+#| warning: false
+#| message: false
+#| eval: true
+
+ggplot(cleaned_par, aes(x = Green.Spaces, y = TOTAL_CRIMES)) +
+  geom_point(alpha = 0.5) + 
+  geom_smooth(method = "lm", se = FALSE) +
+  labs(
+    x = "Total Area of Green Spaces in square kilometers",
+    y = "Total Crimes per 100000",
+    title = "Relationship between Green Spaces and Total Crimes"
+  ) +
+  theme_minimal()
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+my_model <-
+  readRDS(file = here::here("C:/Users/shahj/OneDrive/Documents/greencrimes/models/model.rds"))
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-one
+#| tbl-cap: "Explanatory models of the crime rates based on the green spaces"
+#| warning: false
+
+modelsummary::modelsummary(
+  list(
+    "Model" = my_model
+  ),
+  statistic = "mad",
+  fmt = 2 
+)
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-three
+#| fig-cap: "The number of crimes predicted by the model for the area of green spaces per neighbourhood"
+#| warning: false
+
+plot_predictions(my_model, condition = "Green.Spaces") +
+  labs(x = "Green Spaces (square kilometers)",
+       y = "Crime Rate (per 100000)") +
+  theme_classic()
+```
 
 # Discussion
+In this study, I analyzed the relationship between green spaces and crime rates in Toronto's neighborhoods using gamma regression models.
+The results suggest a positive relationship between the amount of green spaces and crime rates, indicating that areas with more green spaces tend to have higher crime rates.
+However, it is important to note that this relationship does not imply causation.
 
-## First discussion point {#sec-first-point}
+One of the key insights gained from this study is the potential impact of green spaces on crime rates in urban areas.
+The findings suggest that neighborhoods with more green spaces may face unique challenges in terms of crime prevention and urban planning.
+This knowledge can inform policymakers and urban planners in their efforts to create safer and more sustainable cities.
 
-TBD
+## Weaknesses and Next Steps
 
-## Second discussion point
+Several limitations in this study should be acknowledged.
+Firstly, the green space data used in this analysis is from 2011, which may not accurately reflect the current green space distribution in Toronto.
+Additionally, the process of reducing the number of neighborhoods from 158 to 138 due to the removal the NA and 0 values could introduce bias into the analysis (so a total of 20 nighbourhoods were not considered in this study).
 
-TBD
+To deepen the understanding of relationships between green spaces and crime rates, future research should focus on using more recent green space data and exploring other factors that may influence crime rates in urban areas.
+For instance, economic status, social factor, and other demographics relating to the different neighbourhoods.
+Additionally, conducting similar studies in other cities instead of Toronto could help to determine the general idea behind these findings.
 
-## Third discussion point
-
-TBD
-
-## Weaknesses and next steps
-
-TBD
+## Conclusion
+In conclusion, this study provides details about the relationship between green spaces and crime rates in urban areas.
+By understanding the impact of green spaces on crime rates, the policymakers, government and urban planners can work towards creating even more safe and sustainable cities for their residents.
 
@@
 
-\appendix
+# Appendix
 
-# Appendix {-}
+## Datasheet
 
-# Datasheet
+## Posterior predictive check
+This check is a way to assess the goodness-of-fit of a model by comparing its predictions to actual observations.
+(a) Peaks in @fig-pp check around 1000-1500 indicate areas where the model may not accurately capture the data's variability or patterns.
+(b) The posterior estimates are more in the middle, indicating greater certainty after adding in the data, whereas prior estimates are more dispersed, reflecting greater uncertainty before observing the data.
 
-TBD
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-pp
+#| layout-ncol: 2
+#| fig-cap: "Examining how the model fits and is affected by the data"
+#| fig-subcap: ["Posterior prediction check", "Comparing posterior with prior"]
 
-# Model details {#sec-model-details}
+pp_check(my_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
 
-# Posterior predictive check
+posterior_vs_prior(my_model) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
+```
 
-TBD
+## Diagnostics
+The diagnostics help assess whether the model has sampled from the posterior distribution effectively.
+In @fig-mcmc (a) The trace plot shows that chain 4 appears to have higher variability than the others.
+(b) The Rhat plot displays values around 1 indicating convergence of the chains.
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-mcmc
+#| fig-cap: "Checking the convergence of the MCMC algorithm"
+#| fig-subcap: ["Trace plot", "Rhat plot"]
+#| layout-ncol: 2
 
-# Diagnostics
+plot(my_model, "trace")
 
-TBD
+plot(my_model, "rhat")
+```
 
@@
 
-@citeR
+@citeR @citeODT @burtka2023 @shepley2019 @vilkhovlaw

### 2024-04-19 scripts/01-download_data.R

@@ -5,17 +5,19 @@
 # Contact: jena.shah@mail.utoronto.ca
 # License: MIT
 
-
-#### Workspace setup ####
+#### Set up workspace ####
 library(opendatatoronto)
 library(tidyverse)
 
 #### Download data ####
-# Download data from opendatatoronto (1) wellbeing-toronto-environment (2) neighbourhood-crime-rates.
-# Download (1) as the_green_spaces and (2) as the_crime_rates.
+# We want two datasets. (1) Green Spaces and (2) Crime Rates.
+# Green Spaces dataset: data was manually downloaded from Toronto's Open Data Catalogue (cited in references) as green_data.csv.
+
+# Crime Rates dataset: use the code provided on https://open.toronto.ca/dataset/neighbourhood-crime-rates/.
+package <- show_package("neighbourhood-crime-rates")
+resources <- list_package_resources("neighbourhood-crime-rates")
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+crime_data <- filter(datastore_resources, row_number()==1) %>% get_resource()
 
 #### Save data ####
-write_csv(the_crime_rates, "inputs/data/crime_rates.csv") 
-write_csv(the_green_spaces, "inputs/data/green_spaces.csv")
-
-         
+write.csv(crime_data, "C:/Users/shahj/OneDrive/Documents/greencrimes/data/raw/crime_data.csv", row.names = FALSE)

### 2024-04-19 scripts/02-data_cleaning.R

@@ -5,11 +5,36 @@
 # Contact: jena.shah@mail.utoronto.ca
 # License: MIT
 
-#### Workspace setup ####
-library(tidyverse)
+#### Set up workspace ####
+library(dplyr)
+library(arrow)
+
+#### Merge raw data ####
+# Start off with reading raw data.
+crime_data <- read.csv("C:/Users/shahj/OneDrive/Documents/greencrimes/data/raw/crime_data.csv")
+green_data <- read.csv("C:/Users/shahj/OneDrive/Documents/greencrimes/data/raw/green_data.csv")
+
+# We want to combine crime_data's HOOD_ID and green_data's Neighbourhood.Id.
+combined_data <- merge(crime_data, green_data, by.x = "HOOD_ID", by.y = "Neighbourhood.Id")
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+# Filter out variables being used.
+keep_cols <- c("HOOD_ID", "AREA_NAME", "ASSAULT_RATE_2023", "AUTOTHEFT_RATE_2023",
+               "BREAKENTER_RATE_2023", "ROBBERY_RATE_2023", "THEFTOVER_RATE_2023", 
+               "Green.Spaces")
 
+filtered_data <- combined_data %>% 
+  select(all_of(keep_cols))
+
+# Add a column for total crimes per neighborhood.
+filtered_data <- combined_data %>% 
+  select(all_of(keep_cols)) %>%
+  mutate(
+    TOTAL_CRIMES = ASSAULT_RATE_2023 + AUTOTHEFT_RATE_2023 + BREAKENTER_RATE_2023 + ROBBERY_RATE_2023 + THEFTOVER_RATE_2023,
+    TOTAL_CRIMES = replace_na(TOTAL_CRIMES, 0)
+  )
 #### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+# Save as a csv.
+write.csv(filtered_data, "C:/Users/shahj/OneDrive/Documents/greencrimes/data/cleaned/cleaned_data.csv", row.names=FALSE)
+# Save as a parquet.
+write_parquet(filtered_data, "C:/Users/shahj/OneDrive/Documents/greencrimes/data/cleaned/cleaned_data.parquet")

### 2024-04-19 scripts/03-test_data.R

@@ -6,7 +6,29 @@
 # License: MIT
 
 
-#### Workspace setup ####
+#### Set up workspace ####
 library(tidyverse)
 
 #### Test data ####
+# Start off with reading the csv.
+cleaned_data <- read.csv("C:/Users/shahj/OneDrive/Documents/greencrimes/data/cleaned/cleaned_data.csv")
+
+# Check if all crime rate variables are between 0 and 100000 (since they are the number of crimes per 100000).
+all_cols <- c("ASSAULT_RATE_2023", "AUTOTHEFT_RATE_2023", "BREAKENTER_RATE_2023",
+              "ROBBERY_RATE_2023", "THEFTOVER_RATE_2023", "TOTAL_CRIMES")
+between_0_and_100000_check <- sapply(cleaned_data[all_cols], function(x) all(x >= 0 & x <= 100000, na.rm = TRUE))
+
+# Check if the Green.Spaces column values are 0 or more.
+green_spaces_check <- all(cleaned_data$Green.Spaces >= 0)
+
+# Check if the HOOD_ID values are from 1 to 140.
+hood_id_check <- all(cleaned_data$HOOD_ID %in% 1:140)
+
+# Check if the AREA_NAME values are characters.
+area_name_check <- all(sapply(cleaned_data$AREA_NAME, is.character))
+
+# Combine all three tests.
+all_checks <- all(between_0_and_100000_check, green_spaces_check, hood_id_check, area_name_check)
+
+# Print result.
+print(all_checks)

### 2024-04-19 scripts/04-model.R

@@ -5,21 +5,25 @@
 # Contact: jena.shah@mail.utoronto.ca
 # License: MIT
 
-
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 library(rstanarm)
 
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+#### MAKE MODEL ####
+# Read the cleaned data.
+cleaned_par <- read_parquet("C:/Users/shahj/OneDrive/Documents/greencrimes/data/cleaned/cleaned_data.parquet") %>%
+  filter(!is.na(TOTAL_CRIMES) & TOTAL_CRIMES != 0 & !is.na(Green.Spaces) & Green.Spaces != 0)
 
-### Model data ####
-# Model here assigned to my_model
-
-#### Save model ####
-saveRDS(
-  my_model,
-  file = "models/my_model.rds"
+# Fit the gamma regression model.
+model_total_crimes <- stan_glm(
+  TOTAL_CRIMES ~ Green.Spaces,
+  data = cleaned_par,
+  family = Gamma,
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  seed = 123
 )
 
-
+# Save the model.
+saveRDS(model_total_crimes, "C:/Users/shahj/OneDrive/Documents/greencrimes/models/model.rds")

### 2024-04-19 scripts/00-simulate_data.R
Commit

@@ -1,12 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates data.
-# Author: Jena Shah
-# Date: 02 April 2024
-# Contact: jena.shah@mail.utoronto.ca
-# License: MIT
-
-
-#### Workspace setup ####
-library(tidyverse)
-
-#### Simulate data ####

### 2024-04-19 other/datasheet/datasheet_template.qmd
Commit

@@
 1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
-    - TBD
+    - The dataset was created to analyze the relationship between the amount of green spaces in Toronto's neighborhoods and the corresponding crime rates.
+The goal was to understand if there is a correlation between green spaces and crime rates, which could help urban planners and policymakers in enhancing urban green spaces to potentially reduce crime rates.
 2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
-    - TBD
+    - Jena Shah created the dataset. She is a student at the University of Toronto.
 3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
-    - TBD
-4. *Any other comments?*
-    - TBD
+    - It was free.
 

### 2024-04-19 paper/paper.qmd
Commit

@@
 date-format: long
-abstract: "This study examines the relationship between green spaces and crime rates in Toronto's neighborhoods using gamma regression models.
-The analysis reveals a positive association between the amount of green spaces and crime rates, indicating that neighborhoods with more green spaces tend to have higher crime rates.
+abstract: "This study examines the relationship between green spaces and crime rates in Toronto's neighbourhoods using gamma regression models.
+The analysis reveals a positive association between the amount of green spaces and crime rates, indicating that neighbourhoods with more green spaces tend to have higher crime rates.
 This finding emphasizes the relation between urban green spaces and public safety, suggesting that the presence of green spaces may present challenges for crime prevention and urban planning.
@@
 1.  HOOD_ID (Neighbourhood ID): This variable serves as a unique
-    identifier for each neighborhood in Toronto, allowing us to
+    identifier for each neighbourhood in Toronto, allowing us to
     differentiate between different areas in our analysis.
@@
 2.  AREA_NAME (Neighbourhood Name): This variable provides the name of
-    each neighborhood, providing context and a more intuitive
+    each neighbourhood, providing context and a more intuitive
     understanding of the data.
@@
 3.  Green.Spaces (Amount of Green Spaces): This variable quantifies the
-    amount of green spaces in each neighborhood, measured in square
+    amount of green spaces in each neighbourhood, measured in square
     kilometers. Green spaces include parks, forests, and other areas
@@
     This variable represents the total number of reported crimes per
-    neighborhood, standardized to a rate per 100000 people. It provides
-    a measure of the overall crime burden in each neighborhood,
+    neighbourhood, standardized to a rate per 100000 people. It provides
+    a measure of the overall crime burden in each neighbourhood,
     accounting for differences in population size.
@@
 objective of examining the relationship between green spaces and crime
-rates. The inclusion of the neighborhood ID and name allows us to
-identify and differentiate between neighborhoods, while the Green.Spaces
+rates. The inclusion of the neighbourhood ID and name allows us to
+identify and differentiate between neighbourhoods, while the Green.Spaces
 variable provides a measure of the presence and extent of green spaces
@@
 overall crime rates, allowing us to assess the impact of green spaces on
-crime in different neighborhoods. @fig-one shows the different variables
+crime in different neighbourhoods. @fig-one shows the different variables
 and we can see the right-skew in the crimes total, which can help us
@@
 This study was done to investigate the relationship between the abundance of green
-spaces and crime rates in urban neighborhoods in Toronto by using a
+spaces and crime rates in urban neighbourhoods in Toronto by using a
 gamma regression model. This approach is suitable for analyzing count
@@
 
--   Dependent Variable: TOTAL_CRIMES (total crimes per neighborhood per
+-   Dependent Variable: TOTAL_CRIMES (total crimes per neighbourhood per
     100000 people).
@@
 relationship between green spaces and crime rates in Toronto's
-neighborhoods, providing a statistically interpretable
+neighbourhoods, providing a statistically interpretable
 approach to examining this important issue.
@@
 
-In the results, I found that green spaces have a statistically significant relationship with crime rates in Toronto's neighborhoods.
-The gamma regression model showed that as the amount of green spaces in a neighborhood increases, the total crimes per 100000 people also tend to increase.
+In the results, I found that green spaces have a statistically significant relationship with crime rates in Toronto's neighbourhoods.
+The gamma regression model showed that as the amount of green spaces in a neighbourhood increases, the total crimes per 100000 people also tend to increase.
 Specifically, for every unit increase in green spaces (measured in square kilometers), there was a corresponding increase in total crimes.
@@
 It shows a concentration of data points at low to medium crime rates and low levels of green spaces.
-As the amount of green spaces increases, there are much fewer data points and they are more scattered, indicating higher crime rates in neighborhoods with more green spaces.
+As the amount of green spaces increases, there are much fewer data points and they are more scattered, indicating higher crime rates in neighbourhoods with more green spaces.
 The line of best fit in the scatterplot shows a nearly straight line, yet still highlighting the positive relationship between green spaces and crime rates.
@@
 # Discussion
-In this study, I analyzed the relationship between green spaces and crime rates in Toronto's neighborhoods using gamma regression models.
+In this study, I analyzed the relationship between green spaces and crime rates in Toronto's neighbourhoods using gamma regression models.
 The results suggest a positive relationship between the amount of green spaces and crime rates, indicating that areas with more green spaces tend to have higher crime rates.
@@
 One of the key insights gained from this study is the potential impact of green spaces on crime rates in urban areas.
-The findings suggest that neighborhoods with more green spaces may face unique challenges in terms of crime prevention and urban planning.
+The findings suggest that neighbourhoods with more green spaces may face unique challenges in terms of crime prevention and urban planning.
 This knowledge can inform policymakers and urban planners in their efforts to create safer and more sustainable cities.
@@
 Firstly, the green space data used in this analysis is from 2011, which may not accurately reflect the current green space distribution in Toronto.
-Additionally, the process of reducing the number of neighborhoods from 158 to 138 due to the removal the NA and 0 values could introduce bias into the analysis (so a total of 20 nighbourhoods were not considered in this study).
+Additionally, the process of reducing the number of neighbourhoods from 158 to 138 due to the removal the NA and 0 values could introduce bias into the analysis (so a total of 20 nighbourhoods were not considered in this study).
 

### 2024-04-19 other/datasheet/datasheet_template.qmd
Commit

@@
 Please provide a description.*
-	- TBD
+	- Green spaces area in square kilometers and crime rates out of 100000 people.
 2. *How many instances are there in total (of each type, if appropriate)?*
-	- TBD
+	- 138 of each type, based on neighbourhoods in Toronto.
 3.
@@
 7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
-	- TBD
+	- The relationshio between
 8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
-	- TBD
+	- No.
 9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
-	- TBD
+	- Yes, there are 158 neighbourhoods in Toronto but the dataset has 138 due to removal of na or 0 values.
 10.
@@
 Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
-	- TBD
+	- It is self-contained.
 11.
@@
 If so, please provide a description.*
-	- TBD
+	- No.
 12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
-	- TBD
+	- Crime rates may cause anxiety.
 13.
@@
 If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
-	- TBD
+	- No.
 14.
@@
 If so, please describe how.*
-	- TBD
+	- No.
 15.
@@
 If so, please provide a description.*
-	- TBD
-16. *Any other comments?*
-	- TBD
+	- No.
 
@@
 	- TBD
-12. *Any other comments?*
-	- TBD
 
@@
 3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
-	- TBD
-4. *Any other comments?*
-	- TBD
+	- R is cited.
 
@@
 5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
-	- TBD
-6. *Any other comments?*
 	- TBD
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
-	- TBD
-7. *Any other comments?*
 	- TBD
@@
 	- TBD
-8. *Any other comments?*
-	- TBD
 
